### 2018-03-19 CO749-GraphColoring/assignments/A4.tex
question 3 preamble

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%
%END MACROS
%Page style
\pagestyle{fancy}

@@

	Let $G$ be a graph and $F \subset E(G)$ such that $G-F$ is planar. If for every two distinct edges $e_1, e_2 \in F$ we have $d(e_1, e_2) \geq d$ then $G$ is $5$-list-colorable.
\end{theorem}
\begin{proof}
Choose $d = 4r^2+2r+1$ where $r$ is as in Theorem \ref{th:closeness}.
\paragraph{}
Suppose the theorem is false. Let $(G,F)$ be a minimal counterexample with respect to $|V(G)| + |E(G)| + |F|$.
\paragraph{}
Label the elements of $F$ as $\{e_1=u_1v_1, \dots, e_k = u_kv_k\}$. For each $e_i \in F$ let $P_i$ be a $u_i-v_i$ path of length at most $2r$ in $G-F$, provided such a path exists, and let $P_i$ be the empty graph otherwise. Let $S\subset G$ be the graph consisting of $V(F)$ and $\bigcup_{i=1}^k P_i$.
\paragraph{}
Hence we claim $S$ consists of a disjoint union of path components $P_i$ for $i \in \{1,\dots, k\}$ as well as disjoint vertex components $u_j$, $v_j$ where there is no $u_j-v_j$ path of length at most $2r$ in $G-F$. To see this, suppose $P_i$ and $P_j$ intersect at a vertex $x$ for $i\neq j$. Without loss of generality say $d(u_i,x) \leq d(v_i,x)$ and $d(u_j, x) \leq d(v_j, x)$. Then $d(u_i, x) \leq r$ and $d(u_j,x) \leq r$. So we have
$$d(u,i,u,j) \leq r + r = 2r < d$$
a contradiction. The argument follows similarly if a disjoint vertex component intersects a path component.
\paragraph{}
Now we claim that $S$ has at least two components. By our previous argument it suffices to show $|F| \geq 2$. Indeed by Theorem \ref{th:g-e-choosable} if $|F| \leq 1$ then $G$ is $5$-list-colorable, a contradiction.
\paragraph{}
Now we claim $G$ is $S$-critical. Let $H$ be a proper subgraph of $G$ containing $S$. By minimality of $G$, since $H-F$ is planar, and the distances between vertices in $H$ is at least their distance in $G$, $H$ is $5$-list-colorable. Let $\phi$ be a $5$-list-coloring of $H$. If $\phi$ extends to $G$ then $G$ is $5$-list-colorable, a contradiction. Hence $\phi$ does not extend. Thus $G$ is $S$-critical.
\paragraph{}
In order to apply Theorem \ref{th:closeness} with $G-F$ and $S$ it remains to verify $G-F$ is connected. Suppose $G-F$ is disconnected. Then there exist  $u_i,v_i \in F$ such that $u_i$ and $v_i$ lie in distinct components of $G-F$. Re-embed on the plane, if necessary, the components of $G-F$ containing $u_i$ and $v_i$ so $u_i$ and $v_i$ lie incident to the outer face. Then join $u_i$ and $v_i$ by their edge $e_i$ drawn as a curve in the outer face. Hence $G - (F\backslash \{e_i\})$ is planar. But then $(G, F\backslash\{e_i\})$ is a smaller counterexample than $(G,F)$ contradicting minimality.
\paragraph{}
By Theorem \ref{th:closeness} there exist distinct components $S_1$ and $S_2$ such that there is in $G-F$ is
$$d(S_1,S_2) \leq r(|V(S_1)| + |V(S_2)|).$$
First consider the case that $S_1 = u_i$ and $S_2=v_j$ are single vertex components. Then we have
$$d(u,v) \leq r(1+1) = 2r.$$
So if $i=j$ this contradicts that $u$ and $v$ are not joined by a path in $S$, and if $i\neq j$ this contradicts that
$$d(e_i,e_j) \geq d > 2r.$$
Now consider the case that $S_1 =P_i$ and $S_2 = P_j$ are path components. Then there exist $x_i$ on $P_i$ and $x_j$ on $P_j$ such that
$$d(x_i, x_j) \leq r(|V(S_1)| + |V(S_2)|) \leq r(2r + 2r) = 4r^2.$$
Without loss of generality say $d(u_i,x_i) \leq d(v_i,x_i)$ and $d(u_j, x_i) \leq d(v_j, x_i)$. Then $d(u_i, x_i) \leq r$ and $d(u_j,x_i) \leq r$. So we have
$$d(u_i,u_j) \leq r + 4r^2 + r = 4r^2 + 2r < d$$
contradicting that $d(e_i,e_j) \geq d.$ The argument where $S_1$ is a path component and $S_2$ is a vertex component follows similarly.
\paragraph{}
Therefore in any case we obtain a contradiction, as desired.
\end{proof}
\end{document}

>>>>

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
\DeclarePairedDelimiter\norm{\lVert}{\rVert}%

\DeclareMathOperator{\defi}{def}
%END MACROS
%Page style
\pagestyle{fancy}

@@

	Let $G$ be a graph and $F \subset E(G)$ such that $G-F$ is planar. If for every two distinct edges $e_1, e_2 \in F$ we have $d(e_1, e_2) \geq d$ then $G$ is $5$-list-colorable.
\end{theorem}
\begin{proof}
Choose $d = 4r^2+2r+1$ where $r$ is as in Theorem \ref{th:closeness}.
\paragraph{}
Suppose the theorem is false. Let $(G,F)$ be a minimal counterexample with respect to $|V(G)| + |E(G)| + |F|$.
\paragraph{}
Label the elements of $F$ as $\{e_1=u_1v_1, \dots, e_k = u_kv_k\}$. For each $e_i \in F$ let $P_i$ be a $u_i-v_i$ path of length at most $2r$ in $G-F$, provided such a path exists, and let $P_i$ be the empty graph otherwise. Let $S\subset G$ be the graph consisting of $V(F)$ and $\bigcup_{i=1}^k P_i$.
\paragraph{}
Hence we claim $S$ consists of a disjoint union of path components $P_i$ for $i \in \{1,\dots, k\}$ as well as disjoint vertex components $u_j$, $v_j$ where there is no $u_j-v_j$ path of length at most $2r$ in $G-F$. To see this, suppose $P_i$ and $P_j$ intersect at a vertex $x$ for $i\neq j$. Without loss of generality say $d(u_i,x) \leq d(v_i,x)$ and $d(u_j, x) \leq d(v_j, x)$. Then $d(u_i, x) \leq r$ and $d(u_j,x) \leq r$. So we have
$$d(u,i,u,j) \leq r + r = 2r < d$$
a contradiction. The argument follows similarly if a disjoint vertex component intersects a path component.
\paragraph{}
Now we claim that $S$ has at least two components. By our previous argument it suffices to show $|F| \geq 2$. Indeed by Theorem \ref{th:g-e-choosable} if $|F| \leq 1$ then $G$ is $5$-list-colorable, a contradiction.
\paragraph{}
Now we claim $G$ is $S$-critical. Let $H$ be a proper subgraph of $G$ containing $S$. By minimality of $G$, since $H-F$ is planar, and the distances between vertices in $H$ is at least their distance in $G$, $H$ is $5$-list-colorable. Let $\phi$ be a $5$-list-coloring of $H$. If $\phi$ extends to $G$ then $G$ is $5$-list-colorable, a contradiction. Hence $\phi$ does not extend. Thus $G$ is $S$-critical.
\paragraph{}
In order to apply Theorem \ref{th:closeness} with $G-F$ and $S$ it remains to verify $G-F$ is connected. Suppose $G-F$ is disconnected. Then there exist  $u_i,v_i \in F$ such that $u_i$ and $v_i$ lie in distinct components of $G-F$. Re-embed on the plane, if necessary, the components of $G-F$ containing $u_i$ and $v_i$ so $u_i$ and $v_i$ lie incident to the outer face. Then join $u_i$ and $v_i$ by their edge $e_i$ drawn as a curve in the outer face. Hence $G - (F\backslash \{e_i\})$ is planar. But then $(G, F\backslash\{e_i\})$ is a smaller counterexample than $(G,F)$ contradicting minimality.
\paragraph{}
By Theorem \ref{th:closeness} there exist distinct components $S_1$ and $S_2$ such that there is in $G-F$ is
$$d(S_1,S_2) \leq r(|V(S_1)| + |V(S_2)|).$$
First consider the case that $S_1 = u_i$ and $S_2=v_j$ are single vertex components. Then we have
$$d(u,v) \leq r(1+1) = 2r.$$
So if $i=j$ this contradicts that $u$ and $v$ are not joined by a path in $S$, and if $i\neq j$ this contradicts that
$$d(e_i,e_j) \geq d > 2r.$$
Now consider the case that $S_1 =P_i$ and $S_2 = P_j$ are path components. Then there exist $x_i$ on $P_i$ and $x_j$ on $P_j$ such that
$$d(x_i, x_j) \leq r(|V(S_1)| + |V(S_2)|) \leq r(2r + 2r) = 4r^2.$$
Without loss of generality say $d(u_i,x_i) \leq d(v_i,x_i)$ and $d(u_j, x_i) \leq d(v_j, x_i)$. Then $d(u_i, x_i) \leq r$ and $d(u_j,x_i) \leq r$. So we have
$$d(u_i,u_j) \leq r + 4r^2 + r = 4r^2 + 2r < d$$
contradicting that $d(e_i,e_j) \geq d.$ The argument where $S_1$ is a path component and $S_2$ is a vertex component follows similarly.
\paragraph{}
Therefore in any case we obtain a contradiction, as desired.
\end{proof}

\section{}
We may use the following theorem.
\begin{theorem}\label{th:def}
	There exists $c>0$ such that the following holds:

	Let $G$ be a plane graph, let $S$ be a connected subgraph of $G$, and let $L$ be a $5$-list-assignment for $G$. If $G$ is $S$-critical with respect to $L$ then
	$$v(G|S) \leq c(\defi(G|S) - 1),$$
	where $\defi(G|S) = e(G|S) - 3v(G|S).$
\end{theorem}
We will prove the following lemma, under the assumption that a minimum counterexample $G$ has a vertex in $V(G)\backslash V(S)$ with at least three neighbours in $S$.
\begin{lemma}
	Let $G$ be a plane graph with connected subgraph $S$ and let $L$ be a $5$-list-assignment for $G$. If $\phi$ is an $L$-coloring of $S$ that extends to an $L$-coloring of $G$ then
	$$\log_2 E(\phi) \geq (v(G|S) - (c+1)\defi(G|S))/9,$$
	where $E(\phi)$ is the number of extensions of $\phi$ to $G$ and $c$ is the constant of Theorem \ref{th:def}.
\end{lemma}
\begin{proof}
	Suppose not. Let $(G,S,L)$ be a minimum counterexample, minimizing $v(G|S)$. We may assume $G$ has a vertex $v$ in $V(G)\backslash V(S)$ with at least three neighbours in $S$.
	\paragraph{}

\end{proof}
\end{document}
